---
title: "simpleroptions for R package developers"
author: "Luca Cerone"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{simpleroptions-usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

`simpleroptions` implements a simple framework to manage option files in your packages and scripts,
although I think is in packages were it is the most useful.

```{r, include=FALSE}
unlink("~/.simpleroptions_vignette")
```

To use it is very simple, in your package you create an `OptionsManager` object (or more) 
that you can export and document as a normal function.

To create an `OptionManager` is as easy as:

```{r}
library(simpleroptions)
pkgOpt <- OptionsManager("~/.simpleroptions_vignette/example.json",
                        default_options = list(option_1 = "value_1",
                                               option_2 = 1234)
                        )
```

When the code is run the first time `simpleroptions` takes care of creating the directories that are needed
and creates the file `example.json` in the directory `~/.simpleroptions_vignette`, containing the list `default_options` stored as a json object. The permissions of the newly created files are set by default to
"600" (read and write for the user who created only), although this only applies to Linux and Mac, in Windows
this isn't true. You can change the permissions of the file using `permissions = ` when you call `OptionsManager`.

The next time pkgOpt is instantiated, simpleroptions realizes that the file `~/.simpleroptions_vignette/example.json`
already exists and reads the options values from the file rather than using the values provided in `default_options`.

`pkgOpt` is an R6 object of class OptionsManager that has mainly two methods: `get` and `set`.

As the names imply, `get` is used to get the option values, `set` is used to set the value of an option. 
By default when `set` is called the value is also changed in the json file (if you want to disable this
use `auto_save = FALSE` when you call `OptionsManager`).

You can use `get` for the default value of your function arguments, for example 

```{r}
myfun <- function(x = pkgOpt$get("option_1")) {
  print(x)
}

myfun()
```

If you try to get a value that is not defined an error is thown:

```{r, error = TRUE}
pkgOpt$get("option_3")
```

If you export the `pkgOpt` object, your users will be able to change the options to their desired values using
`set`, for example:

```{r}
pkgOpt$set(option_1 = "new_value_1", option_2 = 9999)
```

You can see that the values have been changed:
```{r}
myfun()
```

and in the file too:

```{r}
cat(readr::read_file("~/.simpleroptions_vignette/example.json"))
```

To see the values currently in use you can access 

```{r}
pkgOpt$current_options
```

Note that by default a user can set options that you didn't specify in default_options, for example

```{r}
pkgOpt$set(option_3 = "unknown_option")
pkgOpt$current_options
```

If you want to prevent this, you can use the option `strict = TRUE` when creating the `ObjectManager`.

We can simulate what would happen in that case:

```{r, error = TRUE}
pkgOpt$strict <- TRUE
pkgOpt$set(option_3 = "new_unknown")
```

The main reason `strict` is not enabled by default is because as you develop your package you might want to 
add new options. With the current implementation when the user upgrades to the new version, `simpleroptions`
would throw an error, which is why at the moment `strict = FALSE` the default behaviour.
